\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-06-27 coqimpl.tex
Reply some of the comments.

 Final Field a' : nat := 1.
 Field c : a' = 1 := eq_refl.
 Field bop : nat -> nat -> nat := (+).
 Fail Field comm_add : ...

@@

\textbf{Reasoning Requires Non-Overridable.}
This compilation strategy
has already allowed fields to be type-dependent on former fields thus fulfilling
\ref{langdesign-req1}. However, the current implementation is not satisfactory enough. There are cases where we need to expose
the concrete definition of a given field (especially when we are proving
properties about it).

@@

This example might sound trivial and contrived. Let's look at the
example field "add".\YZ{This 'add' field is no longer there.} Now that we only expose the type of "add", we
cannot prove its commutativity in "comm_add" (because we don't have
access to its concrete definition) but we can override "add" with
subtraction function in "Family I2". This makes sense because we
shouldn't have (inherited) commutativity property for subtraction
anyway. This is contrary to how "a'" is declared "Final", the second "c"
can prove that it's equal to 1 but "a'" is not overridable in the
"Family I2".

@@

\textbf{Special Fields: ``Extensible'' Inductive Type and Defining Recursor.}
We don't actually support extensible inductive types from
\ref{langdesign-req2} in Coq (because that require non-trivial research
effort on theory side---i.e. {What is the semantics of extensible
inductive types?}---and engineering),\YZ{I do consider that your plugin supports extensible inductive types. It's just that they are implemented as a plugin, rather than in the Coq core.}
but we simulate extensible
inductive type using overriding---when extending an inductive type $A$
with new constructors $c$, our plugin will generate a new inductive
definition $A'$ with this new $c$ on the fly and feed it to Coq.  But
still, we must be careful on the exposing data in the family type (the
compiled module type) for the following field.

@@

To construct a recursor as in \ref{langdesign-req3}, we\YZ{A source of confusion in this section is that the reader often has to guess who performs the actions: is it the programmer or the Coq plugin?} create for each
constructor a handler and aggregate them into a family. The whole
family "neg_handler" can be considered as a large matching case. 
Then we use
"FRecursor" to specify the motive and inductive type. Here "FRecursor"
will feed Coq
a module "v_33_34" for doing type-checking---what it does is basically
assert the existence of the recursor and apply to see if handlers can
pass the type checking by Coq.

@@

\textbf{Propositional Partial Recursor.}
Apparently, the exposed module "b_3" in \cref{fig:plugin-example2} is
not expressive enough to really fulfill \ref{langdesign-req2}. For
example, we cannot prove the constructors are injective, i.e.,
\mintinline{Coq}{self_B.tt ≠ self_B.ff}.\YZ{Is it really impossible to prove constructor injectivity without partial recursors and with recursors only? My confusion after reading these paras is that I still don't know why partial recursors are needed.}
In the original setting, to
prove injectivity of constructors of a (vanilla) inductive type, we only
require a recursor and its computational rules. However, as we pointed
out earlier, the recursor of an inductive type cannot be exposed into
the context for the sake of extensibility. Hence, we need to alter the
formulation of the recursor: we need to have a better
characterization of ``extensible inductive types'' exported in the
compiled module type as a part of the interface for the following fields
to use.

>>>>

 Final Field a' : nat := 1.
 Field c : a' = 1 := eq_refl.
 Field bop : nat -> nat -> nat := (+).
 Fail Field comm_bop : ...

@@

\textbf{Reasoning Needs Non-Overridable.}
This compilation strategy
has already allowed fields to be type-dependent on former fields thus fulfilling
\ref{langdesign-req1}. However, the current implementation is not satisfactory enough. There are cases where we need to expose
the concrete definition of a given field (especially when we are proving
properties about it).

@@

This example might sound trivial and contrived. Let's look at the
example field binary operation "bop" initially set to be addition.\YZ{This 'add' field is no longer there.}\EDJreply{Thanks. Fixed} Now that we only expose the type of "bop", we
cannot prove its commutativity in "comm_bop" (because we don't have
access to its concrete definition) but we can override "bop" with
subtraction function in "Family I2". This makes sense because we
shouldn't (have the ability to) inherit commutativity property for subtraction
anyway. This is contrary to how "a'" is declared "Final", the second "c"
can prove that it's equal to 1 but "a'" is not overridable in the
"Family I2".

@@

\textbf{Special Fields: ``Extensible'' Inductive Type and Defining Recursor.}
We support ``extensible'' inductive types from
\ref{langdesign-req2} in Coq to some extent\footnote{Because to give a satisfactory formal definition and semantic for extensible inductive type is not something this paper aims at, and obviously requires more theoretical research effort},\YZ{I do consider that your plugin supports extensible inductive types. It's just that they are implemented as a plugin, rather than in the Coq core.}\EDJreply{I rephrased it. It is more like I don't think this is really the final form of extensible inductive type that everyone is happy about}
by simulating extensible
inductive type using overriding---when extending an inductive type $A$
with new constructors $c$, our plugin will generate a new inductive
definition $A'$ with this new $c$ on the fly and feed it to Coq.  But
still, we must be careful on the exposing data in the family type (the
compiled module type) for the following field.

@@

To construct a recursor as in \ref{langdesign-req3}, we\YZ{A source of confusion in this section is that the reader often has to guess who performs the actions: is it the programmer or the Coq plugin?}\EDJreply{Good idea. Let me reorgnaize this section to ``first programming interface'', then ``plugin implementation''} create for each
constructor a handler and aggregate them into a family. The whole
family "neg_handler" can be considered as a large matching case. 
Then we use
"FRecursor" to specify the motive and inductive type. Here "FRecursor"
will feed Coq
a module "v_33_34" for doing type-checking---what it does is basically
assert the existence of the recursor and apply to see if handlers can
pass the type checking by Coq.

@@

\textbf{Propositional Partial Recursor.}
Apparently, the exposed module "b_3" in \cref{fig:plugin-example2} is
not expressive enough to really fulfill \ref{langdesign-req2}. For
example, we cannot prove the constructors are injective, i.e.,
\mintinline{Coq}{self_B.tt ≠ self_B.ff} because from the perspective of Coq they are just two terms declared in the module type.\YZ{Is it really impossible to prove constructor injectivity without partial recursors and with recursors only? My confusion after reading these paras is that I still don't know why partial recursors are needed.}\EDJreply{The issue is exactly, at that point of programming, for extensibility, there is no access to the concrete recursor. Concrete recursors also cannot be postulated as part of the interface because it breaks extensibility.}
In the original setting, to
prove injectivity of constructors of a (vanilla) inductive type, we only
require a recursor and its computational rules. However, as we pointed
out earlier, the recursor of an inductive type cannot be exposed into
the context for the sake of extensibility. Hence, we need to alter the
formulation of the recursor: we need to have a better
characterization of ``extensible inductive types'' exported in the
compiled module type as a part of the interface for the following fields
to use.
